### 2023-08-09 chapters/introduction.tex
Finalize model-agnostic paper and add stroke paper

Historically, uncertainty estimation has been a central topic in the field of statistics. Uncertainty has traditionally been addressed through methods like confidence intervals, hypothesis testing, and Bayesian inference. Confidence intervals provide a range of plausible values for a parameter, while hypothesis testing evaluates the significance of differences between groups or variables \cite{blitzstein_introduction_2019}. Bayesian inference, on the other hand, offers a powerful framework to incorporate prior knowledge and update beliefs based on observed data using probability theory. These approaches have been extensively studied and applied in various domains, from social sciences to natural sciences, to improve the robustness and generalizability of statistical analyses \cite{gelman_bayesian_2013}. However, the widespread adoption in diverse applications of machine learning, and in particular deep neural networks, has posed new challenges in dealing with uncertainty.

Machine learning algorithms are designed to learn patterns from data and make predictions or decisions based on those patterns. Since many applications of machine learning are concerned with real-world phenomena, it is important to understand the uncertainty associated with the predictions made by the model. For instance, in medical diagnosis, knowing the uncertainty in a model's prediction is crucial when the consequences of a false positive or false negative can be significant. In speech recognition, knowing that a certain word was hard to understand for the model can help avoid misinterpretation of the transcribed text.

@@

The research in this thesis covers a varied set of topics. One group of the produced studies are concerned with generative latent variable models and their applications to uncertainty estimation and speech modelling \cite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. Another group of studies are concerned with self-supervised speech representation learning and automatic speech recognition \cite{borgholt_scaling_2021,borgholt_we_2021,mohamed_selfsupervised_2022,borgholt_brief_2022}. A final group

@@

covering both representation learning and uncertainty estimation and their applications to speech recognition and automated medical coding.

>>>>

Uncertainty estimation has been a central topic in the field of statistics. It has traditionally been addressed through methods like confidence intervals, hypothesis testing, and Bayesian inference. Confidence intervals provide a range of plausible values for a parameter, while hypothesis testing evaluates the significance of differences between groups or variables \cite{blitzstein_introduction_2019}. Bayesian inference, on the other hand, offers a powerful framework to incorporate prior knowledge and update beliefs based on observed data using probability theory. These approaches have been extensively studied and applied in various domains, from social sciences to natural sciences, to improve the robustness and generalizability of statistical analyses \cite{gelman_bayesian_2013}. However, the widespread adoption in diverse applications of machine learning, and in particular deep neural networks, has posed new challenges in dealing with uncertainty.

Machine learning algorithms are designed to learn patterns from data and make predictions or decisions based on those patterns. Since many applications of machine learning are concerned with real-world phenomena, it is important to understand the uncertainty associated with the predictions made by such a model. For instance, in medical diagnosis, knowing the uncertainty in a model's prediction is crucial when the consequences of a false positive or false negative can be significant. In speech recognition, knowing that a certain word was hard to understand for the model can help avoid misinterpretation of the transcribed text.

@@

The research done over the course of the project at hand covers a varied set of topics. One group of the produced studies is concerned with generative latent variable models and their applications to uncertainty estimation and speech modelling \cite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. Another group of studies are concerned with self-supervised speech representation learning and automatic speech recognition \cite{borgholt_scaling_2021,borgholt_we_2021,mohamed_selfsupervised_2022,borgholt_brief_2022}. A final group deals with applications within the medical domain including recognition of stroke cases in calls to medical helplines \cite{wenstrup_retrospective_2023} and medical coding of clinical notes \cite{edin_automated_2023}.

@@

### 2023-08-09 chapters/paper_benchmarking.tex
Finalize model-agnostic paper and add stroke paper

Several works have applied LVMs to speech. Among the first contributions were the VRNN \cite{chung_recurrent_2015} and SRNN \cite{fraccaro_sequential_2016} which can be seen as conditional VAEs per timestep. Other recent LVMs include the FH-VAE \cite{hsu_unsupervised_2017}, which leverages an additional latent variable to capture global features, and Z-forcing \cite{goyal_z-forcing_2017}, which resembles the SRNN but includes an auxiliary task in the latent space to increase its utilization. The VQ-VAE \cite{oord_neural_2018} is a hybrid between an LVM and an autoregressive model which uses a quantized latent space to improve the quality of generated samples. The Stochastic WaveNet \cite{lai_stochastic_2018} and STCN \cite{aksan_stcn_2019} use WaveNet encoder and decoders and temporally independent latent variables.

>>>>

Several works have applied LVMs to speech. Among the first contributions were the VRNN \cite{chung_recurrent_2015} and SRNN \cite{fraccaro_sequential_2016} which can be seen as conditional VAEs per timestep. Other recent LVMs include the FH-VAE \cite{hsu_unsupervised_2017}, which leverages an additional latent variable to capture global features, and Z-forcing \cite{goyal_zforcing_2017}, which resembles the SRNN but includes an auxiliary task in the latent space to increase its utilization. The VQ-VAE \cite{oord_neural_2018} is a hybrid between an LVM and an autoregressive model which uses a quantized latent space to improve the quality of generated samples. The Stochastic WaveNet \cite{lai_stochastic_2018} and STCN \cite{aksan_stcn_2019} use WaveNet encoder and decoders and temporally independent latent variables.

### 2023-08-09 chapters/paper_benchmarking_workshop.tex
Conditional skipping of compilation
Revise skipping
Finalize model-agnostic paper and add stroke paper

\chapter[benchmarking generative latent variable models for speech]{Benchmarking Generative Latent Variable Models for Speech}
\label{chp:benchmarking}

@@

\iffalse

@@

\paragraph{Scope and related work} 
At a high level, this benchmark brings order to LVM model comparisons for speech and also provides useful reference implementations of the models\footnote{\texttt{ \href{https://github.com/JakobHavtorn/benchmarking-lvms}{github.com/JakobHavtorn/benchmarking-lvms}}}. 
An extended version of this paper is available at \href{https://arxiv.org/abs/2202.12707}{arXiv:2202.12707}.
LVMs based on the VAE are of interest due to their ability to learn an approximate posterior distribution over latent variables. This makes them useful beyond generation for e.g. semi-supervised learning \cite{kingma_semi-supervised_2014} and anomaly detection \cite{havtorn_hierarchical_2021}. 
In this paper, we focus on the VRNN \cite{chung_recurrent_2015}, SRNN \cite{fraccaro_sequential_2016} and STCN \cite{aksan_stcn_2019} since they are similar to the original VAE framework. 
We exclude some related work from the benchmark. Specifically, the FH-VAE \cite{hsu_unsupervised_2017} adds a global latent variable but we exclude it due to it's segmentation of the training data. We also exclude Z-forcing \cite{goyal_z-forcing_2017} which resembles the SRNN, but deviates from maximum likelihood training by using an auxiliary loss. We exclude the VQ-VAE \cite{oord_neural_2018} which is a hybrid LVM and autoregressive model, since it uses a quantized latent space. Finally, we exclude the Stochastic WaveNet \cite{lai_stochastic_2018} which is similar, but inferior, to the STCN \cite{aksan_stcn_2019}.

@@

\fi

>>>>

\chapter[benchmarking generative latent variable models for speech]{Benchmarking Generative Latent Variable Models for Speech}
\label{chp:paper-benchmarking-workshop}
\ifthenelse{\equal{\skippapers}{true}}{}{

@@

@@

\paragraph{Scope and related work} 
At a high level, this benchmark brings order to LVM model comparisons for speech and also provides useful reference implementations of the models\footnote{\texttt{ \href{https://github.com/JakobHavtorn/benchmarking-lvms}{github.com/JakobHavtorn/benchmarking-lvms}}}. 
An extended version of this paper is available at \href{https://arxiv.org/abs/2202.12707}{arXiv:2202.12707}.
LVMs based on the VAE are of interest due to their ability to learn an approximate posterior distribution over latent variables. This makes them useful beyond generation for e.g. semi-supervised learning \cite{kingma_semi-supervised_2014} and anomaly detection \cite{havtorn_hierarchical_2021}. 
In this paper, we focus on the VRNN \cite{chung_recurrent_2015}, SRNN \cite{fraccaro_sequential_2016} and STCN \cite{aksan_stcn_2019} since they are similar to the original VAE framework. 
We exclude some related work from the benchmark. Specifically, the FH-VAE \cite{hsu_unsupervised_2017} adds a global latent variable but we exclude it due to it's segmentation of the training data. We also exclude Z-forcing \cite{goyal_zforcing_2017} which resembles the SRNN, but deviates from maximum likelihood training by using an auxiliary loss. We exclude the VQ-VAE \cite{oord_neural_2018} which is a hybrid LVM and autoregressive model, since it uses a quantized latent space. Finally, we exclude the Stochastic WaveNet \cite{lai_stochastic_2018} which is similar, but inferior, to the STCN \cite{aksan_stcn_2019}.

@@

}

### 2023-08-09 chapters/paper_review/04_benchmarking.tex
Finalize model-agnostic paper and add stroke paper

  \scriptsize

@@

  \resizebox*{0.99\textwidth}{!}{%

@@

    \multirow{3}{*}{MockingJay \cite{liu_mockingjay_2020}} & \multirow{3}{*}{\makecell[l]{LS 360 hrs}} & PC & LS 360 hrs & LS test-clean & - & 0.36\textsuperscript{\ref{footnote:low-resource}}/1.8\textsuperscript{\ref{footnote:low-resource}}/3.6\textsuperscript{\ref{footnote:low-resource}}/18\textsuperscript{\ref{footnote:low-resource}}/45\textsuperscript{\ref{footnote:low-resource}}/360  hrs \\ \cline{3-7}

@@

    \multirow{2}{*}[0mm]{DeCoAR \cite{ling_deep_2020}} & \multirow{2}{*}[0mm]{\makecell[l]{LS 100/360/460/\\\:\:\:960 hrs,\\ WSJ si284}} & ASR & WSJ si284 & WSJ eval92 & - & \makecell[l]{25\textsuperscript{\ref{footnote:low-resource}}/40\textsuperscript{\ref{footnote:low-resource}}/81 hrs} \\ \cline{3-7}

@@

    \makecell[l]{vq-wav2vec + \\\:DiscreteBERT \cite{baevski_effectiveness_2020}} & LS 960 hrs & ASR & LS 100 hrs & \makecell[l]{LS test-clean,\\ LS test-other} & - & \makecell[l]{10 mins\textsuperscript{\ref{footnote:low-resource}},\\ 1\textsuperscript{\ref{footnote:low-resource}}/10\textsuperscript{\ref{footnote:low-resource}}/100 hrs} \\ \hline

@@

    HuBERT \cite{hsu_hubert_2021} & \makecell[l]{LS 960 hrs,\\ LL 60k hrs} & ASR & LS 960 hrs & \makecell[l]{LS test-clean\\ LS test-other} & - & \makecell[l]{10 mins\textsuperscript{\ref{footnote:low-resource}},\\ 1\textsuperscript{\ref{footnote:low-resource}}/10\textsuperscript{\ref{footnote:low-resource}}/100/960 hrs} \\ \hline

@@

    UniSpeech-SAT \cite{chen_unispeechsat_2021} & \makecell[l]{LL 60k hrs\\\;\;\;+ GigaSpeech-10k\\\;\;\;+ VP-24k} & Multi & SUPERB & SUPERB & \checkmark & See SUPERB \cite{yang_superb_2021} paper for details \\ \hline
    WavLM \cite{chen_wavlm_2021} & \makecell[l]{LL 60k hrs\\\;\;\;+ GigaSpeech-10k\\\;\;\;+ VP-24k} & Multi & SUPERB & SUPERB & \checkmark & See \cite{yang_superb_2021} for details \\ \hline
    \multirow{4}{*}[0mm]{XLS-R \cite{babu_xlsr_2021}} & \multirow{4}{*}[0mm]{\makecell[l]{VP-400k + MLS\\\;\;\;+ CV-dataset-7k + VL\\\;\;\;+ BBL}} & ASR & VP, MLS, CV-dataset, BBL, LS & VP, MLS, CV-dataset, BBL, LS & - & \multirow{4}{*}[0mm]{See \cite{babu_xlsr_2021} for details} \\ \cline{3-6}

@@

\begin{table}[ht]
  \centering
  \footnotesize
  \caption{Tasks where the state of the art is models with SSL pre-training.}
  \label{table:sota_performance}
  \renewcommand*\arraystretch{1.2}
  \begin{tabular}{llllll}  
    \toprule
    Tasks & Dataset & non-SSL & SSL \\
    \midrule
    ASR (WER $\downarrow$) & LS test-clean/other & 2.1/4.0 \cite{xu_iterative_2020} & 1.4/2.6 \cite{zhang_pushing_2020} \\ \hline
    IC (Acc $\uparrow$) & FSC & 98.8 \cite{lugosch_speech_2019} & 99.3\cite{chen_unispeechsat_2021} \\ \hline
    SID (Acc $\uparrow$) & VoxCeleb1 & 94.8 \cite{hajibabaei_unified_2018} & 95.5 \cite{chen_wavlm_2021} \\ \hline
    ASV (EER $\downarrow$) & VoxCeleb1 & 3.1 \cite{hajavi_siamese_2021} & 2.4 \cite{wang_finetuned_2021} \\ \hline
    QbE (MTWV $\uparrow$) & QUESST (EN) & 10.6 \cite{rodriguez-fuentes_gttsehu_2014} & 11.2\cite{chen_unispeechsat_2021} \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

  \tiny

@@

  \resizebox*{0.97\textwidth}{!}{%

@@

    \multirow{3}{*}{MockingJay \cite{liu_mockingjay_2020}} & \multirow{3}{*}{\makecell[l]{LS 360 hrs}} & PC & LS 360 hrs & LS test-clean & - & 0.36\textsuperscript{\ref{footnote:low-resource}} /1.8\textsuperscript{\ref{footnote:low-resource}} /3.6\textsuperscript{\ref{footnote:low-resource}} /18\textsuperscript{\ref{footnote:low-resource}} /45\textsuperscript{\ref{footnote:low-resource}} /360  hrs \\ \cline{3-7}

@@

    \multirow{2}{*}[0mm]{DeCoAR \cite{ling_deep_2020}} & \multirow{2}{*}[0mm]{\makecell[l]{LS 100/360/\\\:\:\:460/960 hrs,\\ WSJ si284}} & ASR & WSJ si284 & WSJ eval92 & - & \makecell[l]{25\textsuperscript{\ref{footnote:low-resource}}/40\textsuperscript{\ref{footnote:low-resource}}/81 hrs} \\ \cline{3-7}

@@

    \makecell[l]{vq-wav2vec + \\DiscreteBERT \cite{baevski_effectiveness_2020}} & LS 960 hrs & ASR & LS 100 hrs & \makecell[l]{LS test-clean,\\ LS test-other} & - & \makecell[l]{10 mins\textsuperscript{\ref{footnote:low-resource}},\\ 1\textsuperscript{\ref{footnote:low-resource}}/10\textsuperscript{\ref{footnote:low-resource}}/100 hrs} \\ \hline

@@

    HuBERT \cite{hsu_hubert_2021} & \makecell[l]{LS 960 hrs,\\ LL 60k hrs} & ASR & LS 960 hrs & \makecell[l]{LS test-clean\\ LS test-other} & - & \makecell[l]{10 mins\textsuperscript{\ref{footnote:low-resource}}, 1\textsuperscript{\ref{footnote:low-resource}}\\/10\textsuperscript{\ref{footnote:low-resource}}/100/960 hrs} \\ \hline

@@

    UniSpeech-SAT \cite{chen_unispeechsat_2021} & \makecell[l]{LL 60k hrs\\\;\;+ GigaSpeech-10k\\\;\;+ VP-24k} & Multi & SUPERB & SUPERB & \checkmark & See SUPERB \cite{yang_superb_2021} paper for details \\ \hline
    WavLM \cite{chen_wavlm_2021} & \makecell[l]{LL 60k hrs\\\;\;+ GigaSpeech-10k\\\;\;+ VP-24k} & Multi & SUPERB & SUPERB & \checkmark & See \cite{yang_superb_2021} for details \\ \hline
    \multirow{4}{*}[0mm]{XLS-R \cite{babu_xlsr_2021}} & \multirow{4}{*}[0mm]{\makecell[l]{VP-400k + MLS\\\;\;+ CV-dataset-7k\\\;\;+ VL\\\;\;+ BBL}} & ASR & VP, MLS, CV-dataset, BBL, LS & VP, MLS, CV-dataset, BBL, LS & - & \multirow{4}{*}[0mm]{See \cite{babu_xlsr_2021} for details} \\ \cline{3-6}

@@

\begin{table}[t!]
  \centering
  \footnotesize
  \caption{Tasks where the state of the art is models with SSL pre-training.}
  \label{table:sota_performance}
  \renewcommand*\arraystretch{1.2}
  \begin{tabular}{llllll}  
    \toprule
    Tasks & Dataset & non-SSL & SSL \\
    \midrule
    ASR (WER $\downarrow$) & LS test-clean/other & 2.1/4.0 \cite{xu_iterative_2020} & 1.4/2.6 \cite{zhang_pushing_2020} \\ \hline
    IC (Acc $\uparrow$) & FSC & 98.8 \cite{lugosch_speech_2019} & 99.3\cite{chen_unispeechsat_2021} \\ \hline
    SID (Acc $\uparrow$) & VoxCeleb1 & 94.8 \cite{hajibabaei_unified_2018} & 95.5 \cite{chen_wavlm_2021} \\ \hline
    ASV (EER $\downarrow$) & VoxCeleb1 & 3.1 \cite{hajavi_siamese_2021} & 2.4 \cite{wang_finetuned_2021} \\ \hline
    QbE (MTWV $\uparrow$) & QUESST (EN) & 10.6 \cite{rodriguez-fuentes_gttsehu_2014} & 11.2\cite{chen_unispeechsat_2021} \\
    \bottomrule
  \end{tabular}
\end{table}
